---
title: "Homework 1"
author: "Karthika Shankar"
date: "2026-09-26"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains information on penguins, including species, island, bill length, bill depth, flipper length, body mass, sex, and year. The species in the dataset include Adelie, Chinstrap, and Gentoo penguins, and the islands include Biscoe, Dream, and Torgersen. The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

```{r penguins_scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

ggsave("penguins_scatterplot.png")
```

# Problem 2

```{r problem2_dataframe}
set.seed(1234)
problem2_df = tibble(
  random_sample = rnorm(10),
  positive = random_sample > 0,
  character_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  factor_vector = factor(c("low", "medium", "high", "low", "medium",
                           "high", "low", "medium", "high", "low"))
)

problem2_df
```

```{r problem2_means}
mean(pull(problem2_df, random_sample))
mean(pull(problem2_df, positive))
mean(pull(problem2_df, character_vector))
mean(pull(problem2_df, factor_vector))
```

The mean can be calculated for the numeric and logical variables, but not for the character or factor variables. For the logical variable, R treats `TRUE` as 1 and `FALSE` as 0, so the mean represents the proportion of values that are `TRUE`. The character and factor variables return `NA` because they are not numeric or logical.

```{r problem2_coercion, results = "hide", warning = FALSE}
as.numeric(pull(problem2_df, positive))
as.numeric(pull(problem2_df, character_vector))
as.numeric(pull(problem2_df, factor_vector))
```

Converting the logical variable to numeric changes `FALSE` to 0 and `TRUE` to 1. Converting the character variable to numeric produces `NA` values because the character values can not be interpreted as numbers. Converting the factor variable to numeric gives the integer codes assigned to the factor levels rather than meaningful numerical values. This explains why the mean works for logical variables but not directly for character or factor variables.
